A preliminary constraint grammar for Russian

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60237 72bbbca6-d526-0410-a7d9-f06f51895060

\needspace{6\baselineskip} % keeps example all together (no orphan line) 
\ex. Morphosyntactically congruent homographs \label{ex:MScongruent}
\a. \rus{з\'{a}мок} \emph{z\'{a}mok} `castle.\textsubscript{SG-NOM}'\\
	\rus{зам\'{о}к} \emph{zam\'{o}k} `lock.\textsubscript{SG-NOM}'
\b. \rus{з\'{a}мков} \emph{z\'{a}mkov} `castle.\textsubscript{PL-GEN}'\\
	\rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'\\
    etc. \ldots

\rr{is this clear that the paradigm continues, but we don't show it? One of the reviewers thought the example was incomplete}

@@

    \rr{R2: The first two rules in 6.2 are a bit odd - the effect would be exactly the same without the first context bracket. Did you mean 0C instead of 0?}

@@

As can be seen from Figure~\ref{fig:curve}, incorporating the constraint grammar 
improves the performance of the HMM tagger, an improvement of nearly 5\% in accuracy,
similar to that reported by \newcite{hulden12} for the same amount of training data. 
In Figure~\ref{fig:curve}, it appears that the HMM alone is much more dependent on 
training corpus size than the voting setup, which improves very little between a training
corpus size of 5,000 and 9,000. This seems to indicate that further improvements depend
more on the CG than on training data.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.70\textwidth]{graphics/learning-curve.pdf}
  \caption{Learning curve for three taggers, \texttt{hunpos} with no lexicon, 
    \texttt{hunpos} with a lexicon, and \texttt{hunpos} with a lexicon and the 
    Russian constraint grammar in a voting set up.}
  \label{fig:curve}
\end{figure*}
\rr{R2: If I interpret Fig. 3 correctly, the HMM system alone is much more dependent on training corpus size than the voting set-up, where accuracies actually appear to become asymptotic already with 8-10,000 words in the training corpus, indicating that further improvement depends more on the CG than the HMM-component.}

>>>>

\needspace{6\baselineskip} % keeps example all together (no orphan line) 
\ex. Morphosyntactically congruent homographs \label{ex:MScongruent}
\a. \rus{з\'{a}мок} \emph{z\'{a}mok} `castle.\textsubscript{SG-NOM}'\\
	\rus{зам\'{о}к} \emph{zam\'{o}k} `lock.\textsubscript{SG-NOM}'
\b. \rus{з\'{a}мков} \emph{z\'{a}mkov} `castle.\textsubscript{PL-GEN}'\\
	\rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'\\
    etc.

@@

@@

As can be seen from Figure~\ref{fig:curve}, incorporating the constraint grammar 
improves the performance of the HMM tagger, an improvement of nearly 5\% in accuracy,
similar to that reported by \newcite{hulden12} for the same amount of training data. 
In Figure~\ref{fig:curve}, it appears that the HMM alone is much more dependent on 
training corpus size than the voting setup, which improves very little between a training
corpus size of 5,000 and 9,000.

@@

\begin{figure*}
  \centering
  \includegraphics[width=0.70\textwidth]{graphics/learning-curve.pdf}
  \caption{Learning curve for three taggers, \texttt{hunpos} with no lexicon, 
    \texttt{hunpos} with a lexicon, and \texttt{hunpos} with a lexicon and the 
    Russian constraint grammar in a voting set up.}
  \label{fig:curve}
\end{figure*}
